---
title: "High Performance Computing Week 1 Assignment"
author: "Daipayan Bera"
date: "01-31-2025"
output: html_notebook
---
```{r}
#Clearing environment
rm(list = ls())
library(foreach)
library(doParallel) #Importing doParallel package
```

**Q1: Importing required library to import data**

```{r}
library(MASS)
data("Boston")
registerDoSEQ() #Verifying that we are working with serially
```

**Q1: Creating 100 Bootstrap samples of the dataset with the SAME dimension using REPLACE=TRUE**

```{r}
n <- nrow(Boston)
Rprof("serial_sampling")
set.seed("143")
boston_bootstrap <- foreach(i=1:100) %do% {Boston[sample(n, n, replace = TRUE), ]}
Rprof(NULL)
summaryRprof('serial_sampling')
```
*So, here I have used FOREACH function with %DO% to recieve a list of dataframes after bootstraping, by using the SAMPLE function. I have also used RPROF to see how proficient is the segment in terms of time complexity.*

**Q1: Fitting GLM model on each bootstrap sample**

```{r}
Rprof("serial_model")
boston_bootstrap.lm <- foreach(i=1:100) %do% {glm(medv~., data=boston_bootstrap[[i]])}
Rprof(NULL)
summaryRprof("serial_model")
```

*By using the same principle, I have recieved a list of the all fitted model. I have utlized all independent variables as predictor and MEDV as outcome variable.*

**Q2: Retrieving model statistics for all fitted models**

```{r}
if (!requireNamespace("broom", quietly = TRUE)) {
  install.packages("broom")
}
library(broom)
```

*Here, I wanted to receive the model statistics as a dataframe which further could be used for plotting. So I took help of BROOM package (https://cran.r-project.org/web/packages/broom/vignettes/broom.html). It has three main functions, 1. Tidy, 2. Augment, 3. Glance. Here, I am going to use GLANCE. It provides one row summary of all model performance statistics that we are interested in.*

*But the issue is that, if we apply this as function on the list, it is going to provide 100 statistic dataframes of each model. As we want a consolidated dataframe for all model statistics, therefore, I am going to use RBIND with DO.APPLY on the list of dataframes (Returned from GLANCE function)*

```{r}
Rprof("serial_statistics")
boston_boot.lm.fit <- foreach(i=1:100) %do% {
  glance(boston_bootstrap.lm[[i]])}
Rprof(NULL)
summaryRprof("serial_statistics")
```

*Applying RBIND on the list of one-row dataframes*

```{r}
bootstrap_statistics <- do.call(rbind, boston_boot.lm.fit)
```


```{r}
#Lets check the data that we have recieved
head(bootstrap_statistics)
```

**Q3: Now, I was asked to "Aggregate the model fit statistics across the 100 bootstrapped samples and plot the results". So I am choosing MEAN as aggregation.**

```{r}
bootstrap_statistics.mean <- data.frame(apply(bootstrap_statistics,2, mean))
bootstrap_statistics.mean
```

*But, here I ran into a problem, as I typecasted the resultant vector as dataframe. It always showed the statistic names as index or something. No matter what I did to change/remove with my ability. It didn't get affected at all. Also, when I used it to plot, it also didn't appear in the axis. So, I went ahead and crated a separate column with the same entries.*

```{r}
bootstrap_statistics.mean["Statistics"] = c("null.deviance","df.null","logLik","AIC","BIC","deviance","df.residual","nobs")
names(bootstrap_statistics.mean)[1] = "Mean"
bootstrap_statistics.mean
```
*Although, it's still there, but now I have what I need to create a bar plot of the statistic.*

```{r}
library(ggplot2)
serial.plot <-ggplot(data = bootstrap_statistics.mean, aes(x = Statistics, y = Mean)) +
  geom_col() + #Using geom_bar gave stat_count() error, cause for geom_bar, y variable takes a categorical variable as input
  labs(title = "Serially computed bootstrap samples : GLM model statistics(Mean)")
serial.plot
```
**Q3: Computing all other model statistics**

```{r}
summary(bootstrap_statistics)
```
**Q4. Creating bootstrap samples parallelly with foreach function**

```{r}
cl <- makeCluster(detectCores(), type='PSOCK')
registerDoParallel(cl)
getDoParWorkers() #Verifying that we are computing parallelly
```
```{r}
Rprof("parallel_sampling")
set.seed(143)
parallel_boot.samples <- foreach (i=1:100, .packages = c("doParallel", "foreach")) %dopar% {Boston[sample(n, n, replace = TRUE),]}
Rprof(NULL)
summaryRprof("parallel_sampling")
```

**Fitting GLM model on each bootstrap sample (Parallelly)**

```{r}
Rprof("Parallel_model")
parallel_boot.models <- foreach(i=1:100, parallel_boot.samples, .packages = c("doParallel", "foreach")) %dopar% {
  glm(medv~.,data=parallel_boot.samples[[i]]) 
}
Rprof(NULL)
summaryRprof("Parallel_model")
```

**Q4. Retrieving model statistics for all bootstrap models (Using Parallelization)**

```{r}
Rprof("parallel_statistics")
parallel_boot.statistics <- foreach(i=1:100, parallel_boot.models, .packages = c("doParallel", "foreach", "broom")) %dopar% {
  glance(parallel_boot.models[[i]])
}
Rprof(NULL)
summaryRprof("parallel_statistics")
```


**Q4. Retrieving a dataframe consisting all model statistics**

```{r}
parallel_boot.statistics.merge <- do.call(rbind, parallel_boot.statistics) #Using rbind to combine all one-row dataframes (recieved from GLANCE function) into one dataframe to better usability 
parallel_boot.statistics.mean <- data.frame(apply(parallel_boot.statistics.merge,2, mean))
#Creating a separate column consisting all the model statistic labels
parallel_boot.statistics.mean["Statistics"] = c("null.deviance","df.null","logLik","AIC","BIC","deviance","df.residual","nobs")
names(parallel_boot.statistics.mean)[1] = "Mean"
parallel_boot.statistics.mean
```

**Q4. Plotting aggregated (Mean) model statistics for both Parallel and Serial computing**

```{r}
ggplot(data = parallel_boot.statistics.mean, aes(x = Statistics, y = Mean))+
  geom_col()+
  labs(title = "Parallelly computed bootstrap samples : GLM model statistics(Mean)")
serial.plot
```
```{r}
#Stopping cluster and parallel computing
stopCluster(cl)
registerDoSEQ()
```

###Profiling through Microbenchmark

```{r}
serial_bootstrap_glmfit <- function(Boston){
  n=nrow(Boston)
  library(broom)
  set.seed("143")
  serial_bootstrap <- foreach(i=1:100) %do% {Boston[sample(n, n, replace = TRUE), ]}
  serialboot_glm <-foreach(i=1:100) %do% {glm(medv~.,data=serial_bootstrap[[i]])}
  serialboot_glm_stat <- foreach(i=1:100) %do% {glance(serialboot_glm[[i]])}
  serial_stats <- do.call(rbind, serialboot_glm_stat)
}
```

```{r}
parallel_bootstrap_glmfit <- function(Boston){
  cl <- makeCluster(detectCores()-1, type='PSOCK')
  doParallel::registerDoParallel()
  n=nrow(Boston)
  set.seed("143")
  parallel_bootstrap <- foreach(i=1:100, .packages = c("foreach","doParallel")) %dopar% {
    Boston[sample(n, n, replace = TRUE), ]}
  parallelboot_glm <-foreach(i=1:100, .packages = c("foreach","doParallel")) %dopar% {
    glm(medv~.,data=parallel_bootstrap[[i]])}
  parallelboot_glm_stat <- foreach(i=1:100,.packages = c("foreach","doParallel","broom")) %dopar% {
    glance(parallelboot_glm[[i]])}
  parallel_stats <- do.call(rbind, parallelboot_glm_stat)
  stopCluster(cl)
  registerDoSEQ()
}
```


```{r}
microbenchmark::microbenchmark(
  'serial_op' = serial_bootstrap_glmfit(Boston),
  'parallel_op' = parallel_bootstrap_glmfit(Boston),
  times=10
)
```

